Ideas cascading into keystrokes -- modelling writing hesitations as finite mixture process

### 2023-12-20 report/manuscript.Rmd
minor fix

To compare the out-of-sample predictive performance of our models we used Pareto smoothed importance-sampling leave-one-out cross-validation [@vehtari2015pareto; @vehtari2017practical]. Predictive performance was estimated as the sum of the expected log predictive density ($\widehat{elpd}$) and compared via its difference $\Delta\widehat{elpd}$ between models. We also summarise the comparison between models as the absolute value of the ratio of the difference $\Delta\widehat{elpd}$ and its standard error -- $\mid\frac{\Delta\widehat{elpd}}{\text{SE}}\mid$, hence the *z*-score of the difference between models [@sivula2020uncertainty]. Similar to other cross-validation techniques, the advantage of using leave-one-out cross-validation is that more complicated models -- models with more parameters -- are penalised to prevent overfit.

@@

table <- purrr::map_dfr(files, ~read_csv(.) %>% 
                          mutate(dataset = .x)) %>% 
  select(dataset, model, contains("elpd_diff"), se_diff) %>%
  arrange(dataset, desc(elpd_diff)) %>% 
  mutate(across(dataset, ~sub(".*/([^/]+)/[^/]+\\..*", "\\1", .)),
         across(where(is.numeric), ~round(., 0)),
         across(where(is.numeric), ~format(., big.mark = ",")),
         across(everything(), ~str_trim(.))) %>%
  filter(dataset != "spl2_shift") %>% 
  transmute(
    Comparison = model,
    elpd_diff_ratio,
     across(dataset, ~recode(., 
                "cato" = "CATO",
                "spl2" = "SPL2", 
#                "spl2_shift" = "SPL2 (shift + C)", 
                "plantra" = "PLanTra",
                "lift" = "LIFT",
                "c2l1" = "C2L1",
                "gunnexp2" = "GUNNEXP2")),
    Description = case_when(str_detect(model, "M2$") ~ "Single distribution log-Gaussian (M2)",
                            str_detect(model, "M3$") ~ "Single distribution log-Gaussian (unequal variance; M3)",
                            str_detect(model, "M4$") ~ "Two distributions mixture of log-Gaussians (constrained; M4)",
                            str_detect(model, "M5$") ~ "Two distributions mixture of log-Gaussians (unconstrained; M5)"),
            across(elpd_diff, ~str_c(., " (", se_diff, ")")))

table_ratios <- table %>% select(dataset, Comparison, elpd_diff_ratio) %>% 
  arrange(Comparison)

@@

apa_table(table, 
      caption = "Model comparisons. Models were compared stepwise from the simples to the most complex model. The comparison is indicated along with the description of the more complex model in each comparison. A negative difference in $\\Delta\\widehat{elpd}$ indicates higher predictive performance for the more complex model.  Standard error is shown in parentheses.",
      align =c("p{1.5cm}", "p{5cm}", rep("r", 6)), 
      escape = FALSE,
      longtable = T,
      font_size = "small",
      note = "$\\Delta\\widehat{elpd}$ = difference in predictive performance -- estimated as expected log pointwise predictive density -- relative to the model with the highest predictive performance in the top row.")
#                col_spanners = list(" " = 1:2,
#                     "\\textbf{CL21}" = 1:5,
#                     "\\textbf{CATO}" = 6:10,
#                     "\\textbf{GUNNEXP2}" = 11:15, 
#                     "\\textbf{LIFT}" = 16:20,
#                     "\\textbf{PLanTra}" = 21:25,
#                     "\\textbf{SPL2}" = 26:30))
#      stub_indents = list(
#                     "\\textbf{CL21}" = 1:5,
#                     "\\textbf{CATO}" = 6:10,
#                     "\\textbf{GUNNEXP2}" = 11:15, 
#                     "\\textbf{LIFT}" = 16:20,
#                    "\\textbf{PLanTra}" = 21:25,
#                     "\\textbf{SPL2}" = 26:30))
#                     "\\textbf{SPL2 (shift + C)}" = 31:35))

>>>>

To compare the out-of-sample predictive performance of our models we used Pareto smoothed importance-sampling leave-one-out cross-validation [@vehtari2015pareto; @vehtari2017practical]. Predictive performance was estimated as the sum of the expected log predictive density ($\widehat{elpd}$) and compared via its difference $\Delta\widehat{elpd}$ between models. We also summarised the comparison between models as the absolute value of the ratio of the difference $\Delta\widehat{elpd}$ and its standard error $\mid\frac{\Delta\widehat{elpd}}{\text{SE}}\mid$, hence the *z*-score of the difference between models [@sivula2020uncertainty]. Similar to other cross-validation techniques, the advantage of using leave-one-out cross-validation is that more complicated models -- models with more parameters -- are penalised to prevent overfit.

@@

table <- purrr::map_dfr(files, ~read_csv(.) %>% 
                          mutate(dataset = .x)) %>% 
  select(dataset, model, contains("elpd_diff"), se_diff) %>%
  arrange(dataset, desc(elpd_diff)) %>% 
  mutate(across(dataset, ~sub(".*/([^/]+)/[^/]+\\..*", "\\1", .)),
         across(where(is.numeric), ~round(., 0)),
         across(where(is.numeric), ~format(., big.mark = ",")),
         across(everything(), ~str_trim(.))) %>%
  filter(dataset != "spl2_shift") %>% 
  transmute(
    Comparison = model,
    elpd_diff_ratio,
     across(dataset, ~recode(., 
                "cato" = "CATO",
                "spl2" = "SPL2", 
#                "spl2_shift" = "SPL2 (shift + C)", 
                "plantra" = "PLanTra",
                "lift" = "LIFT",
                "c2l1" = "C2L1",
                "gunnexp2" = "GUNNEXP2")),
    Distribution = case_when(str_detect(model, "M2$") ~ "Single log-Gaussian (M2)",
                            str_detect(model, "M3$") ~ "Single log-Gaussian (unequal variance; M3)",
                            str_detect(model, "M4$") ~ "Mixture of two log-Gaussians (constrained; M4)",
                            str_detect(model, "M5$") ~ "Mixture of two log-Gaussians (unconstrained; M5)"),
            across(elpd_diff, ~str_c(., " (", se_diff, ")")))

table_ratios <- table %>% select(dataset, Comparison, elpd_diff_ratio)

@@

table %>% 
  arrange(desc(Comparison)) %>% 
  apa_table(caption = "Model comparisons. Models were compared incrementally from the simples to the most complex model. The comparison is indicated along with the distribution of the more complex model in each comparison. A negative difference in $\\Delta\\widehat{elpd}$ indicates higher predictive performance for the more complex model.  Standard error is shown in parentheses.",
      align =c("p{1.5cm}", "p{5cm}", rep("r", 6)), 
      escape = FALSE,
      longtable = T,
      font_size = "footnotesize",
      note = "$\\Delta\\widehat{elpd}$ = difference in predictive performance -- estimated as expected log pointwise predictive density -- relative to the model with the highest predictive performance in the top row. Model details can be found in Appendix \ref{statistical-models}.")
#                col_spanners = list(" " = 1:2,
#                     "\\textbf{CL21}" = 1:5,
#                     "\\textbf{CATO}" = 6:10,
#                     "\\textbf{GUNNEXP2}" = 11:15, 
#                     "\\textbf{LIFT}" = 16:20,
#                     "\\textbf{PLanTra}" = 21:25,
#                     "\\textbf{SPL2}" = 26:30))
#      stub_indents = list(
#                     "\\textbf{CL21}" = 1:5,
#                     "\\textbf{CATO}" = 6:10,
#                     "\\textbf{GUNNEXP2}" = 11:15, 
#                     "\\textbf{LIFT}" = 16:20,
#                    "\\textbf{PLanTra}" = 21:25,
#                     "\\textbf{SPL2}" = 26:30))
#                     "\\textbf{SPL2 (shift + C)}" = 31:35))
